Iterative construction of replicated designs based on Sobol' sequences

### 2016-06-14 crass/crassnew.tex
hopefully last changes

To compute the next $2^m$ points of the sequence, one can infer from \eqref{dig.net.eq.} that
\begin{equation}\label{Sobol_iter_constr}
(x_{i+2^m,j,1},\dots,x_{i+2^m,j,m+1})^\intercal=(x_{i,j,1},\dots,x_{i,j,m},0)^\intercal\oplus {(c_j^{m+1})}^\intercal,\quad i=0,\dots,2^m-1,
\end{equation}
where $c_j^{m+1}$ is the last column of $C_j^{m+1}$.

Sobol' sequences also have good group structure properties. For any $m\geq 1$, the first $2^m$ points of the sequence form an Abelian group under the following operation $\oplus$,
\begin{gather*}
\vx = \left(\sum_{k=1}^m x_{1,k}2^{-k},\dots,\sum_{k=1}^m x_{s,k}2^{-k}\right)^\intercal\; , \qquad \vz = \left(\sum_{k=1}^m z_{1,k}2^{-k},\dots,\sum_{k=1}^m z_{s,k}2^{-k}\right)^\intercal, \\
\vx\oplus\vz := \left(\sum_{k=1}^m (x_{1,k}+z_{1,k}\mod 2)2^{-k},\dots,\sum_{k=1}^m (x_{s,k}+z_{s,k}\mod 2)2^{-k}\right)^\intercal.
\end{gather*}

@@

Initial sets are $B_0=\{\vx_i\}_{i=0}^{2^r-1}$ and ${B'}_0=\{{\vx'}_i\}_{i=0}^{2^r-1}$. Then, one may take $B_\ell=\{\vx_i\oplus\ve_\ell\}_{i=0}^{2^r-1}$ and ${B'}_\ell=\{{\vx'}_i\oplus{\ve'}_\ell\}_{i=0}^{2^r-1}$ such that $\ve_\ell\notin\mathcal{P}_{\ell-1}$ and ${\ve'}_\ell\notin\mathcal{P'}_{\ell-1}$.  Indeed, $B_0$ is a subgroup of $\left(\mathbb{F}_2^{r}\right)^{s}$. Therefore, we can construct cosets $B_\ell$ as $B_0 \oplus \ve$, under the right $\ve \in \left(\mathbb{F}_2^{r}\right)^{s}$. After that, we may also apply linear scramblings \cite{Matousek} using lower triangular matrices $L$ and $L'$ ...... There are exactly $2^{rs}$ different choices of $\ve$ and ${\ve'}$, and $2^{r(r-1)/2}$ choices of $L$ and $L'$. The maximum number of iterations $\ell$ that can be performed equals the number of cosets, $2^{s}$. If this maximum is reached, designs $\mathcal{P}_\ell$ and $\mathcal{P}'_\ell$ both correspond to a full grid design of $2^{rs}$ points.

%Left multiplications by matrices $L$ and $L'$ in \eqref{dig.shift} and \eqref{ibinom.scrambling}  are called linear scramblings \cite{Matousek} while additions by vectors $\ve_j$ and ${\ve'}_j$ are referred as digital shifts. There are exactly $2^r$ different choices of $\ve_j$ and ${\ve'}_j$ and $2^{r(r-1)/2}$ choices of $L$ and $L'$. To guarantee that new points added will fill unexplored regions of the input space, the vectors $\ve_j$ are selected as follows:
%\begin{itemize}
%\item[i)] at iteration $\ell=0$, vectors $\ve_j$ are all equal to the zero of $\mathbb{F}_2^r$,
%\item[ii)] for $\ell \geq1$, we use the following result:
%$(B_0,\oplus)$ is a subgroup of $\left(\mathbb{F}_2^{r}\right)^{s}$. From Lagrange's Theorem, we can construct cosets $B_\ell$ as $B_0 \oplus \ve$, $\ve=(\ve_1,\dots,\ve_s) \in \left(\mathbb{F}_2^{r}\right)^{s}$
%%, which form a partition of $\left(\mathbb{F}_2^{r}\right)^{s}$
%. At iteration $\ell \geq1$, we select $\ve$ such that $B_{\ell}=B_0 \oplus \ve$ is a new coset different from $\{B_0,\dots,B_{\ell-1}\}$. It will be enough to choose $\ve$ such that $\ve\notin\mathcal{P}_{\ell-1}$. %By identification we have $B_{\ell}=B_0 \oplus \ve$ ensuring that the new points added are all different.

@@

\begin{lemma}
Under the additive construction $\mathcal{P}_\ell$ and $\mathcal{P}'_\ell$ are two replicated designs of order $1$, $\ell \geq 0$.
\end{lemma}
\textit{Proof}:For each $\ell\geq 1$, digital shift and scrambling operations carried out in equations (\ref{dig.shift}) and (\ref{ibinom.scrambling}) are bijections from $\mathbb{F}^r_2$ to $\mathbb{F}^r_2$. Hence, same coordinates of $B_\ell$ and ${B'}_\ell$ will contain the same set of $2^r$ values.

>>>>

To compute the next $2^m$ points of the sequence, one can infer from \eqref{dig.net.eq.} that
\begin{equation}\label{Sobol_iter_constr}
(x_{i+2^m,j,1},\dots,x_{i+2^m,j,m+1})^\intercal=(x_{i,j,1},\dots,x_{i,j,m},0)^\intercal\oplus {(c_j^{m+1})}^\intercal,\quad i=0,\dots,2^m-1,
\end{equation}
where $c_j^{m+1}$ is the last column of $C_j^{m+1}$ and $\oplus$ is the addition in $\mathbb{F}_2$.

Sobol' sequences also have good group structure properties. For any $m\geq 1$, the first $2^m$ points of the sequence form an Abelian group under $\oplus$:
\begin{equation}
\begin{gathered}
\vx = \left(\sum_{k=1}^m x_{1,k}2^{-k},\dots,\sum_{k=1}^m x_{s,k}2^{-k}\right)^\intercal\; , \qquad \vz = \left(\sum_{k=1}^m z_{1,k}2^{-k},\dots,\sum_{k=1}^m z_{s,k}2^{-k}\right)^\intercal, \\
\vx\oplus\vz := \left(\sum_{k=1}^m (x_{1,k}+z_{1,k}\mod 2)2^{-k},\dots,\sum_{k=1}^m (x_{s,k}+z_{s,k}\mod 2)2^{-k}\right)^\intercal.
\end{gathered}
\label{group.operation}
\end{equation}

@@

At step $\ell=0$, $B_0=\{\vx_i^{(0)}\}_{i=0}^{2^r-1}$ and ${B'}_0=\{{\vx'}_i^{(0)}\}_{i=0}^{2^r-1}$ are generated as follows:

for each $i=0,\dots,2^r-1$, point $\vx_i^{(0)} = (x_{i,1}^{(0)},\dots,x_{i,s}^{(0)})^\intercal$ and ${\vx'}_i^{(0)} = ({x'}_{i,1}^{(0)},\dots,{x'}_{i,s}^{(0)})^\intercal$ are obtained by linearly transforming $\vx_i$,
\begin{equation}
\label{linear.scrambl}
\begin{aligned}
(x_{i,j,1}^{(0)},\dots,x_{i,j,r}^{(0)})^\intercal &= L (x_{i,j,1},\dots,x_{i,j,r})^\intercal,\\
({x'}_{i,j,1}^{(0)},\dots,{x'}_{i,j,r}^{(0)})^\intercal &= L' (x_{i,j,1},\dots,x_{i,j,r})^\intercal,\qquad j= 1,\dots,s\,
\end{aligned}
\end{equation}
where $ x_{i,j}^{(0)}=\sum_{k = 1}^r x_{i,j,k}^{(0)}2^{-k}$, $ {x'}_{i,j}^{(0)}=\sum_{k = 1}^r {x'}_{i,j,k}^{(0)}2^{-k}$ and $L$, $L'$ are two full rank lower triangular matrices of size $r \times r$ over $\mathbb{F}_2$. Left multiplications by matrices $L$ and $L'$ in \eqref{linear.scrambl} are called linear scramblings \cite{Matousek}.

At step $\ell \geq 1$, using formula (\ref{group.operation}) we construct:
\begin{equation}
B_\ell=\{\vx_i^{(0)}\oplus\ve_\ell\}_{i=0}^{2^r-1}, \qquad {B'}_\ell=\{{\vx'}_i^{(0)}\oplus{\ve'}_\ell\}_{i=0}^{2^r-1}
\label{dig.shift}
\end{equation}
where $\ve_\ell \in \left(\mathbb{F}_2^{r}\right)^{s} \setminus \mathcal{P}_{\ell-1}$ and ${\ve'}_\ell \in \left(\mathbb{F}_2^{r}\right)^{s} \setminus \mathcal{P'}_{\ell-1}$. Indeed, $B_0$ is a subgroup of $\left(\mathbb{F}_2^{r}\right)^{s}$. Therefore, we construct $B_\ell=B_0 \oplus \ve_\ell$ as a coset of $B_0$ under the right $\ve_\ell$. The same reasoning applies to ${B'}_0$.  Additions by vectors $\ve_j$ and ${\ve'}_j$ in (\ref{dig.shift}) are referred as digital shifts.

There are exactly $2^{rs}$ different choices of $\ve_\ell$ and ${\ve'}_\ell$, and $2^{r(r-1)/2}$ choices of $L$ and $L'$. The maximum number of iterations $\ell$ that can be performed equals the number of cosets, $2^{r(s-1)}$. If this maximum is reached, designs $\mathcal{P}_\ell$ and $\mathcal{P}'_\ell$ both correspond to a full grid design of $2^{rs}$ points.

\bigskip
%Left multiplications by matrices $L$ and $L'$ in \eqref{dig.shift} and \eqref{ibinom.scrambling}  are called linear scramblings \cite{Matousek} while additions by vectors $\ve_j$ and ${\ve'}_j$ are referred as digital shifts. There are exactly $2^r$ different choices of $\ve_j$ and ${\ve'}_j$ and $2^{r(r-1)/2}$ choices of $L$ and $L'$. To guarantee that new points added will fill unexplored regions of the input space, the vectors $\ve_j$ are selected as follows:
%\begin{itemize}
%\item[i)] at iteration $\ell=0$, vectors $\ve_j$ are all equal to the zero of $\mathbb{F}_2^r$,
%\item[ii)] for $\ell \geq1$, we use the following result:
%$(B_0,\oplus)$ is a subgroup of $\left(\mathbb{F}_2^{r}\right)^{s}$. From Lagrange's Theorem, we can construct cosets $B_\ell$ as $B_0 \oplus \ve$, $\ve=(\ve_1,\dots,\ve_s) \in \left(\mathbb{F}_2^{r}\right)^{s}$
%%, which form a partition of $\left(\mathbb{F}_2^{r}\right)^{s}$
%. At iteration $\ell \geq1$, we select $\ve$ such that $B_{\ell}=B_0 \oplus \ve$ is a new coset different from $\{B_0,\dots,B_{\ell-1}\}$. It will be enough to choose $\ve$ such that $\ve\notin\mathcal{P}_{\ell-1}$. %By identification we have $B_{\ell}=B_0 \oplus \ve$ ensuring that the new points added are all different.

@@

\begin{lemma}
Under the additive construction $\mathcal{P}_\ell$ and $\mathcal{P}'_\ell$ are two replicated designs of order $1$, $\ell \geq 0$.
\end{lemma}
\textit{Proof}:For each $\ell\geq 1$, digital shift and scrambling operations carried out in equations (\ref{linear.scrambl}) and (\ref{dig.shift}) are bijections. Hence, same coordinates of $B_\ell$ and ${B'}_\ell$ will contain the same set of $2^r$ values.
